---
title: "Web Mapping With Leaflet"
subtitle: "Team Analytics and Beyond"
format: 
  revealjs:
    slide-number: c/t   
    logo: "https://upload.wikimedia.org/wikipedia/commons/e/e8/Marquette_Golden_Eagles_logo.svg"
date: "`r Sys.Date()`"
author: "**Andrew Hiller, Annie Gibbons, Caroline Kwak, Noah Kaye**"
number-sections: false
from: markdown+emoji
editor: 
  source
---

```{r}
#| label: setup
#| include: false

####################################################
## !!!DO NOT MAKE ANY CHANGE OF THIS CODE CHUNK!!!##
####################################################

# Package names
packages <- c("knitr", "ggplot2", "ggrepel", 
              "tidyverse", "formatR", "dslabs", "janitor", 
              "ggthemes", "plotly", "tidymodels", "kknn")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
    install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
```

## First Slide
Explain what leaflet is

## Second Slide
Side by side of leaflet map versus static map (picture of map?)
Explain why leaflet is better and show some features

## More about why it's better
maybe some slides about situations in which leaflet is better than static map
(using examples from the documentation site)

## Show what we created
end product of marquette map

## Then some slides for the demo
show how we created the map

## Conclusion-ish slide
revisit what was learned in the demo and how it can be useful

## Our data
```{r}
#|echo: false

library(leaflet)

xl_file <- readxl::read_excel("./data/MU Location Coordinates.xlsx")
xl_file

```

##

```{r}

m <- leaflet(data = xl_file) %>%
  addTiles() %>%
  addMarkers(lat = ~Latitude , lng = ~Longitude , popup = ~Type, label = ~Location)

m

getcolor <- function(xl_file){
  sapply(xl_file$Type, function(Type) {
    if(Type == "Residence Hall") {
      "purple"
    }else if(Type == "University Apartment"){
      "orange"
    }else if(Type == "Administrative") {
      "green"
    }else if(Type == "Academic") {
      "blue"
    }else if(Type == "Recreation") {
      "red"
    }else if(Type == "Religious Institution"){
      "yellow"
    }else if(Type == "Parking") {
      "black"
    }else{
      "white"
    }
  })
}

icons <- awesomeIcons(
  icon = 'ios-close',
  iconColor = 'black',
  library = 'ion',
  markerColor = getcolor(xl_file)
)


#leaflet(xl_file) |>
#  addTiles() |>
#  addAwesomeMarkers(~lng, ~lat, icon=icons, label=~as.character(Type))

```

## slide with new icons
```{r}
leaflet(xl_file) |>
  addTiles() |>
  addAwesomeMarkers(lng = ~Longitude, lat = ~Latitude, popup = ~Location, icon=icons,
                    label=~as.character(Type))
```